Reanalysis Step 2: Multiple Imputation

### 2025-08-30 reanalysis-2-multiple_imputation.qmd
fix

# n_eyes can be predictor but not predicted
if ("n_eyes" %in% rownames(pred_matrix)) {
  pred_matrix["n_eyes", ] <- 0  # Not predicted
  # n_eyes can help predict missingness patterns
  # Leave default (1) for using as predictor
}

@@

# Variables that will ACTUALLY be imputed (have missing data AND a method)
vars_to_impute <- final_methods |>
  filter(Will_Be_Imputed) |>
  pull(Variable)
```

@@

# FINAL STEP: Enforce patient-level variable consistency across all datasets
# person_dr, CCI, and DCSI should be consistent within patients
imputed_datasets <- lapply(imputed_datasets, function(imp_data) {
  # Calculate consistent values per patient (using max for all)
  patient_level_vars <- imp_data |>
    group_by(e_mrn_deidentified) |>
    summarise(
      person_dr_consistent = max(person_dr, na.rm = TRUE),
      CCI_consistent = max(CCI, na.rm = TRUE),
      DCSI_consistent = max(DCSI, na.rm = TRUE),
      .groups = 'drop'
    )

>>>>

# n_eyes should not be used as predictor OR predicted
# It's perfectly correlated with fellow variable missingness patterns
if ("n_eyes" %in% rownames(pred_matrix)) {
  pred_matrix["n_eyes", ] <- 0  # Not predicted
  pred_matrix[, "n_eyes"] <- 0  # Not a predictor (causes collinearity)
}

@@

# Variables that will ACTUALLY be imputed (have missing data AND a method)
# Exclude fellow variables and n_eyes even if they have missing data
vars_to_impute <- final_methods |>
  filter(Will_Be_Imputed) |>
  filter(!Variable %in% c("fellow_baseline_VA", "fellow_intermediate_VA", 
                          "fellow_outcome_VA", "n_eyes")) |>
  pull(Variable)
```

@@

# FINAL STEP: Enforce patient-level variable consistency across all datasets
# person_dr, CCI, and DCSI should be consistent within patients
imputed_datasets <- lapply(imputed_datasets, function(imp_data) {
  # Calculate consistent values per patient (using max for all)
  patient_level_vars <- imp_data |>
    group_by(e_mrn_deidentified) |>
    summarise(
      # Handle case where all values are NA (max returns -Inf)
      person_dr_consistent = ifelse(all(is.na(person_dr)), 
                                   NA_real_, 
                                   max(person_dr, na.rm = TRUE)),
      CCI_consistent = ifelse(all(is.na(CCI)), 
                             NA_real_, 
                             max(CCI, na.rm = TRUE)),
      DCSI_consistent = ifelse(all(is.na(DCSI)), 
                              NA_real_, 
                              max(DCSI, na.rm = TRUE)),
      .groups = 'drop'
    )
